Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-05-18 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Added 43.75 w/l to MacRog dehydration data.
MacRog 44 w/l to Table 2.

\begin{table}[htb]
\centering
\caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ Water/lipid molar ratio
$^b$ The number of lipid molecules
$^c$ The number of water molecules
$^d$ Simulation temperature
$^e$ The total simulation time
$^f$ Time frames used in the analysis
$^g$ Reference link for the downloadable simulation files
$^h$ Reference for the full simulation details
}\label{systemsDEHYD}
\begin{tabular}{c c c c c c c c c c}
%\hline
Force field & lipid & $^a$n (w/l)   & $^b$N$_{\rm l}$   &  $^c$N$_{\rm w}$ & $^d$T (K)  & $^e$t$_{{\rm sim}}$(ns)  & $^f$t$_{{\rm anal}}$ (ns)& $^g$Files  &  $^h$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & SI \\
                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
Berger-DLPC-13~\cite{kanduc13}          &   DLPC & 28  &72 & 2016  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc28}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 24  &72 & 1728  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc24}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 20  &72 & 1440  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc20}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 16  &72 & 1152  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc16}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 12  &72 & 864  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc12}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 8  &72 & 576  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc8}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 4  &72 & 288  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc4}]$^*$ & [\citenum{kanduc13}] \\
CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
                                 & POPC   & 31  & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
MacRog\cite{kulig15b}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI  \\
                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & [\citenum{gaffDEHYDfiles}]$^*$ & SI \\
\end{tabular}
\end{table}

@@

\subsubsection{Phospholipid bilayer with low hydration level}
Fig.~\ref{ordPhydr} shows the published~\cite{dvinskikh05b,ulrich94,bechinger91} experimental order parameters
for the glycerol backbone and choline as a function of hydration level.
Despite slight differences in temperature and acyl chain composition,
the three independently reported data sets for the choline ($\beta$ and $\alpha$) segments agree well with each other:
Both order parameters increase with decreasing hydration level.
The glycerol backbone order parameters (g$_{3}$, g$_{2}$, g$_{1}$), in contrast,
have been observed~\cite{dvinskikh05b} to slightly decrease with dehydration.
Note that the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91} measured only 
absolute values, but here
we included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
Consequently, the negative $\beta$ order parameter actually increases with dehydration 
as its absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
\begin{figure}[]
  \centering
  \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_lp.pdf}
%    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
    \caption{\label{ordPhydr}
    The effect of dehydration on glycerol and choline order parameters in experiments.
    The magnitudes of order parameters are measured for DMPC ($^1$H-$^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.
    Note that to elucidate the relative change as a function of hydration level, the simulation results were vertically shifted; 
    the shift magnitudes for each of the force fields are listed (SCH+shift) in the y-label.
  }
\end{figure}

>>>>

\begin{table}[htb]
\centering
\caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ Water/lipid molar ratio
$^b$ The number of lipid molecules
$^c$ The number of water molecules
$^d$ Simulation temperature
$^e$ The total simulation time
$^f$ Time frames used in the analysis
$^g$ Reference link for the downloadable simulation files
$^h$ Reference for the full simulation details
}\label{systemsDEHYD}
\begin{tabular}{c c c c c c c c c c}
%\hline
Force field & lipid & $^a$n (w/l)   & $^b$N$_{\rm l}$   &  $^c$N$_{\rm w}$ & $^d$T (K)  & $^e$t$_{{\rm sim}}$(ns)  & $^f$t$_{{\rm anal}}$ (ns)& $^g$Files  &  $^h$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & SI \\
                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
Berger-DLPC-13~\cite{kanduc13}          &   DLPC & 28  &72 & 2016  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc28}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 24  &72 & 1728  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc24}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 20  &72 & 1440  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc20}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 16  &72 & 1152  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc16}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 12  &72 & 864  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc12}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 8  &72 & 576  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc8}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 4  &72 & 288  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc4}]$^*$ & [\citenum{kanduc13}] \\
CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
                                 & POPC   & 31  & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
MacRog\cite{kulig15b}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                               & POPC   & 44 & 288  & 12600 & 310 & 100 & 80 & [\citenum{macrogFILES}]$^*$ & SI \\
                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI  \\
                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & [\citenum{gaffDEHYDfiles}]$^*$ & SI \\
\end{tabular}
\end{table}

@@

\subsubsection{Phospholipid bilayer with low hydration level}
Fig.~\ref{ordPhydr} shows the published~\cite{dvinskikh05b,ulrich94,bechinger91} experimental order parameters
for the glycerol backbone and choline as a function of hydration level.
Despite slight differences in temperature and acyl chain composition,
the three independently reported data sets for the choline ($\beta$ and $\alpha$) segments agree well with each other:
Both order parameters increase with decreasing hydration level.
The glycerol backbone order parameters (g$_{3}$, g$_{2}$, g$_{1}$), in contrast,
have been observed~\cite{dvinskikh05b} to slightly decrease with dehydration.
Note that the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91} measured only 
absolute values, but here
we included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
Consequently, the negative $\beta$ order parameter actually increases with dehydration 
as its absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
\begin{figure}[]
  \centering
  \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration.pdf}
    \caption{\label{ordPhydr}
    The effect of dehydration on glycerol and choline order parameters in experiments.
    The magnitudes of order parameters are measured for DMPC ($^1$H-$^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.
    Note that to elucidate the relative change as a function of hydration level, the simulation results were vertically shifted; 
    the shift magnitudes for each of the force fields are listed (SCH+shift) in the y-label.
  }
\end{figure}
